---
title: "R Notebook - 3 Data visualisation"
subtitle: "R for Data Science"
output: html_notebook
---

```{r}
#intall.packages("tidyverse")
library(tidyverse)
```

```{r}
#Import data
?ggplot2::mpg
mpg
```
```{r}
str(mpg)
```


```{r}
ggplot(data=mpg) +
  geom_point(mapping = aes(x=displ, y=hwy))
```

3.2.4 Exercises<br>
1. Run ggplot(data = mpg). What do you see?<br>
2. How many rows are in mpg? How many columns?<br>
3. What does the drv variable describe? Read the help for ?mpg to find out.<br>
4. Make a scatterplot of hwy vs cyl.<br>
5. What happens if you make a scatterplot of class vs drv? Why is the plot not useful?<br>

```{r}
#1.
ggplot(data=mpg)

#2.
dim(mpg)

#3.
?mpg
#the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd

#4.
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=cyl, y=hwy))
#Observa-se que carros com menor quantidade de cilindros tendem a apresentar uma maior eficiência.

#5.
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=drv, y=class))
#Um scatter plot dessas features não é muito útil pois ambos os dados são categoricos, apresentados grande overlap sobre seus valores.
```
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy, color=class))
```
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy, size=class))
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy, alpha=class))
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy, shape=class))
```
```{r}
ggplot(data= mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy), color='blue')
```
3.3.1 Exercises
2. Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?<br>
3. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?
4. What happens if you map the same variable to multiple aesthetics?
5. What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)
6. What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y.

```{r}
#2.
str(mpg)
#Observando o emprego de "str" sobre os dados, a maioria dos dados são categóricos, podendo ser representado como feature contínua as variáveis "dipl", "hwy" e "cty".
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y=cty)) +
  geom_point(mapping = aes(x=displ, y=hwy), color='blue')
```

```{r}
#3.
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy, size=cty, color=class))
```
```{r}
ggplot(data = mtcars, aes(x=wt, y=mpg)) + 
  geom_point(shape=21, colour='black', fill='white', size=5, stroke = 5)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy, colour = displ < 5))
```
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy)) + 
  facet_wrap(~ class, nrow=2)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy)) + 
  facet_grid(. ~ class)
```
```{r, fig.height=10}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy)) + 
  facet_grid(class ~ .)
```



3.5.1 Exercises

1. What happens if you facet on a continuous variable?<br>
2. What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?<br>
3. What plots does the following code make? What does . do?<br>
4. Take the first faceted plot in this section:
What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?<br>
5. Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol arguments?<br>
6. When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy)) + 
  facet_grid(. ~ cty)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy)) + 
  geom_smooth(mapping = aes(x=displ, y=hwy), color='green')

ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x=displ, y=hwy))

ggplot(data = mpg, mapping = aes(x=displ, y=hwy)) + 
  geom_point(mapping = aes(color=drv)) + 
  geom_smooth(mapping = aes(linetype = drv, color = drv), se=FALSE)
```

```{r}
ggplot(data = mpg, mapping = aes(x=displ, y=hwy)) + 
  geom_point(mapping = aes(color=drv)) + 
  geom_vline(xintercept = mean(mpg$displ)) + 
  geom_hline(yintercept = mean(mpg$hwy))
```

```{r}
ggplot(data = mpg) + 
  geom_boxplot(mapping = aes(y=hwy, color=drv))
```
```{r}
ggplot(data = mpg, mapping = aes(x=displ, y=hwy)) + 
  geom_point() + 
  geom_smooth(se=FALSE)

ggplot(data = mpg, mapping = aes(x=displ, y=hwy)) + 
  geom_point() + 
  geom_smooth(mapping = aes(group=drv), se=FALSE)
```
```{r}
ggplot(data = mpg, mapping= aes(x=displ, y=hwy)) + 
  geom_point(mapping = aes(color=drv)) + 
  geom_smooth(mapping = aes(color=drv), se=FALSE)

ggplot(data = mpg, mapping = aes(x=displ, y=hwy)) + 
  geom_point(mapping = aes(color=drv)) + 
  geom_smooth(se=FALSE)

ggplot(data = mpg, mapping = aes(x=displ, y=hwy)) + 
  geom_point(mapping = aes(color=drv)) + 
  geom_smooth(mapping = aes(linetype=drv), se=FALSE)

ggplot(data = mpg, mapping = aes(x=displ, y=hwy)) + 
  geom_point(stroke=6, color='white') + 
  geom_point(mapping = aes(color=drv), size=4)
```


```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))
  
```

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x=cut, y=stat(prop), group=1))
```
```{r}
ggplot(data = mpg, mapping = aes(x=class)) + 
  geom_bar(mapping = aes(y=stat(prop), group=1))

ggplot(data = mpg, mapping = aes(x=class)) + 
  geom_bar(mapping = aes(y=stat(prop), group=1, weight = displ)) #https://ggplot2.tidyverse.org/reference/geom_bar.html, weight

ggplot(data = mpg, mapping = aes(x=class)) + 
  geom_bar(mapping = aes(weight = displ))

ggplot(data = mpg, mapping = aes(x=class)) + 
  geom_bar()
```
```{r}
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max,
    fun = median
  )
```

```{r}
ggplot(data = diamonds) + 
  geom_pointrange(mapping = aes(x=cut, y=depth),
                  stat = 'summary',
                  fun.min = min,
                  fun.max = max,
                  fun = median)
  
```


```{r}
ggplot(data = diamonds) + 
  geom_boxplot(mapping = aes(x=cut, y=depth))
```

```{r}
ggplot(data = mtcars, mapping = aes(x= factor(cyl), fill= factor(gear))) + 
  geom_bar(position = position_dodge(width=0.8))
```

```{r}
ggplot(data = mtcars) + 
  geom_boxplot(mapping = aes(y= gear, group=mpg))
```

```{r}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x=mpg, y=wt, size=cyl, color=carb)) + 
  geom_vline(xintercept = median(mtcars$mpg)) + 
  geom_hline(yintercept = median(mtcars$wt)) + 
  facet_wrap(~gear)

mtcars_sc <- as.data.frame(scale(mtcars))

ggplot(data = mtcars_sc) + 
  geom_point(mapping = aes(x=mpg, y=wt, size=cyl, color=carb)) + 
  geom_vline(xintercept = median(mtcars_sc$mpg)) + 
  geom_hline(yintercept = median(mtcars_sc$wt)) + 
  geom_smooth(mapping = aes(x=mpg, y=hp), se=FALSE, color='red') 
 
```


```{r}
ggplot(data = mtcars) + 
  geom_histogram(mapping = aes(x=hp), binwidth = 15)
```

```{r}

ggplot(data = mpg, mapping = aes(x=cty, y=hwy)) + 
  geom_point(position = 'jitter', alpha=0.5) + 
  geom_abline() + 
  labs(title='Hwy vs cty', subtitle = 'Comparação', x='cty', y='hwy', caption = 'Fonte:') + 
  coord_fixed() + 
  geom_smooth(method = 'lm', se= FALSE)

```


```{r}
lm(hwy ~ cty, data=mpg)
```

```{r}
my_variabe_name <- 2.5

my_variabe_n <- 3.5

```

```{r}

```

References<br>
[1]https://r4ds.had.co.nz/data-visualisation.html#creating-a-ggplot
